---
title: "Buoy project"
author: "ziyi bai"
date: "2020/9/25"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document:
    df_print: paged
---
## Introduction: What am I supposed to do in this project?

< p style="font-size:20px">

Buoy is a floating device that can be used for navigation, marker, rescue, research, military purposes. In this project, we will use 20 years weather data collected by one weather buoy near Boston to determine whether these data show us the increasing temperature pattern, which can be known as the problem of global warming. If my model shows me an increasing  trend in temperature, then global warming might be an issue that encourage our concern. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(ggplot2)
library(rstanarm)
library(tidyverse)
library(lubridate)
library(zoo)
```

## Main Body: describe my approach of analysing

< p style="font-size:20px">

The first step in my project is to collect and reorganize data. We collected data from 1987 to 2016, one of my group mate, zijie huang, was responsible for organizing those data. We calculated the average value of ATMP and WTMP for each month and combine them into a new graph, MR_4.

```{r}
## clean data
setwd("/Users/baiziyi/Desktop/MSSP/MA615/buoy/buoy-project")
MR <- read.csv(file="Buoy_MR.csv", header = TRUE)
MR_2 <- as_tibble(MR)
#Change type of Air_tmp and Water_tmp from chr to dbl
MR_2 <- MR_2 %>%
  mutate(ATMP = as.double(ATMP),
         WTMP = as.double(WTMP),
         YYYY_MM_DD = ymd(paste(YYYY,MM,DD,sep = "-"))) %>%
  relocate(YYYY_MM_DD)

#Get rid of tittle line
MR_2 <- filter(MR_2,hh != "hr")

#Get rid of abnormal data
MR_2 <- filter(MR_2,ATMP < 100, WTMP < 100)

#Get daily average Tmp
MR_3_1 <- select(MR_2,-c(2,3,4,7)) %>%
  group_by(YYYY_MM_DD) %>%
  summarize(Avg_ATMP = mean(ATMP))

MR_3_2 <- select(MR_2,-c(2,3,4,6)) %>%
  group_by(YYYY_MM_DD) %>%
  summarize(Avg_WTMP = mean(WTMP))

#Get monthly average Tmp
MR_4_1 <- MR_3_1 %>%
  group_by(month = floor_date(YYYY_MM_DD,"month")) %>%
  summarize(Avg_ATMP = mean(Avg_ATMP))

MR_4_2 <- MR_3_2 %>%
  group_by(month = floor_date(YYYY_MM_DD,"month")) %>%
  summarize(Avg_WTMP = mean(Avg_WTMP))

MR_4 <- inner_join(MR_4_1,MR_4_2)
print(MR_4)
#Export data
write_csv(MR_4,"MR_data_1987_2016.csv")
```

< p style="font-size:20px">

Then, I plotted the data. I plotted the average ATMP and WTMP as y-axis, use month as x-axis using ggplot. It seems that the pattern is messy and it is hard for us to clarify the trend of temperature. Also, we can see that for one year temperature trend, the highest temperature is in July and the lowest is in February. 

```{r}
##Fit model
#Plot data
MR_4 %>%
  ggplot(aes(x = month,y = Avg_ATMP)) + 
  geom_line() +
  geom_point()

MR_4 %>%
  ggplot(aes(x = month,y = Avg_WTMP)) + 
  geom_line() +
  geom_point()

#Check the trend within a year
MR_1988 <- filter(MR_4,year(month)==1988)
MR_1988 %>%
  ggplot(aes(x = month, y = Avg_ATMP)) +
  geom_line() +
  geom_point()
```

< p style="font-size:20px">

Cause these are years long data, so we suppose time series might be helpful. Still, this method just indicated little information about global warming.

```{r}
###Fit time series model
#create TS
TS_MR_4_1 <- ts(select(MR_4_1,Avg_ATMP))

#check trend
plot(TS_MR_4_1,type="b")

#set seasonal index
diff12 = diff(TS_MR_4_1,12)
plot(diff12,type="b")

#check acf pacf
install.packages('astsa', repos = "http://cran.us.r-project.org")
library(astsa)
acf2(diff12,48)

#fit model
sarima(TS_MR_4_1,1,0,0,0,1,1,12)

#predict
sarima.for(TS_MR_4_1,24,1,0,0,0,1,1,12)

#repeat process for water_tmp
TS_MR_4_2 <- ts(select(MR_4_2,Avg_WTMP))
plot(TS_MR_4_2,type="b")
diff12 = diff(TS_MR_4_2,12)
plot(diff12,type="b")
acf2(diff12,48)
sarima(TS_MR_4_2,1,0,0,0,1,1,12)
sarima.for(TS_MR_4_2,24,1,0,0,0,1,1,12)
```

< p style="font-size:20px">

Finally, I decided to make individual graph of each month to show the trend of ATWP and WTMP. For both ATMP and WTMP, the temperature in winter tends to decreasing and the temperature in summer tends to increasing. The plot shows us a nearly linear relationship, so I will build linear regression model following.

```{r}
###Fit lm model

#Red indicates an upward trend
#Blue indicates an downward trend

#plot atmp
par(mfrow=c(3,4))
for(i in 1:12){
  LM_MR_4_1 <- filter(MR_4_1,month(month)==i)
  LM_1 <- lm(Avg_ATMP~month,data = LM_MR_4_1)

  if(coef(LM_1)[2]>0){
    plot(x = LM_MR_4_1$month, y =  LM_MR_4_1$Avg_ATMP,xlab = "month", ylab = "ATMP",main = paste("month",i,sep = "_"))
    abline(coef(LM_1)[1],coef(LM_1)[2], col="RED")
  
  }
  
  if(coef(LM_1)[2]<0){
    plot(x = LM_MR_4_1$month, y =  LM_MR_4_1$Avg_ATMP,xlab = "month", ylab = "ATMP",main = paste("month",i,sep = "_"))
    abline(coef(LM_1)[1],coef(LM_1)[2], col="BLUE")
  
  }
}

#plot water_tmp
par(mfrow=c(3,4))
for(i in 1:12){
  LM_MR_4_2 <- filter(MR_4_2,month(month)==i)
  LM_2 <- lm(Avg_WTMP~month,data = LM_MR_4_2)

  if(coef(LM_2)[2]>0){
    plot(x = LM_MR_4_2$month, y =  LM_MR_4_2$Avg_WTMP,xlab = "month", ylab = "WTMP",main = paste("month",i,sep = "_"))
    abline(coef(LM_2)[1],coef(LM_2)[2], col="RED")
  
  }
  
  if(coef(LM_2)[2]<0){
    plot(x = LM_MR_4_2$month, y =  LM_MR_4_2$Avg_WTMP,xlab = "month", ylab = "WTMP",main = paste("month",i,sep = "_"))
    abline(coef(LM_2)[1],coef(LM_2)[2], col="BLUE")
  
  }
}

```

## Conclusion 

< p style="font-size:20px">

To conclude, 20 or 30 years of data in Boston is still not enough to make more accurate prediction, But as I indicated in the previous chunk, red line means an upward trend of temperature. Those graphs show us that air and water temperatures in Boston tend to increasing in the summer. So, global warming maybe an issue that society needs to take into consideration. 


## Reference

2020. Knitr, Rstudio. Beijing.
2020. ggplot2, Rstudio. Beijing.
2020. rstanarm, Rstudio. Beijing.
2020. tidyverse, Rstudio. Beijing.
2020. lubridate, Rstudio. Beijing.
2020. zoo, Rstudio. Beijing.
2020. astsa, Rstudio. Beijing.